Semilattice Graphs for Distributed Databases

### 2022-09-30 paper/paper.tex
about to refactor a lot!

        \begin{itemize}
            \item A set of keys $K$.
            \item A set of values $V$.
            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key to a finite set of keys, its dependencies.
            \item A join function $j : (\Sigma (k: K, x : V, y : V), d(k, x) \cup d(k, y) \rightarrow V) \rightarrow V$ that joins two values given their dependencies. (The sigma set $\Sigma (a: A), B(a)$ denotes a set of pairs $(a, b)$ where $a \in A$ and $b \in B(a)$.)
        \end{itemize}

@@

        \begin{itemize}
            \item $d$ must be well-founded; that is, there are no infinite sequences $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, \ldots$
            \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g|_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.
            \item $j$ must not grow dependencies; that is, for any $k : K$, $g: K \rightarrow V$, $x, y \in V$, we must have $d(k, x \vee_{k, g} y) \subseteq d(k, x) \cup d(k, y)$.
        \end{itemize}

@@

>>>>

        \begin{itemize}
            \item A set of keys $K$.
            \item A nonempty set of values $V$.
            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key to a finite set of keys, its dependencies.
            \item A join function $j : (\Sigma (k: K, x : V, y : V), d(k, x) \cup d(k, y) \rightarrow V) \rightarrow V$ that joins two values given their dependencies. (The sigma set $\Sigma (a: A), B(a)$ denotes a set of pairs $(a, b)$ where $a \in A$ and $b \in B(a)$.)
        \end{itemize}

@@

        \begin{itemize}
            \item $d$ must be well-founded; that is, there are no infinite dependency chains $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, 2, \ldots$
            \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g|_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.
            \item $j$ must not grow dependencies; that is, for any $k : K$, $g: K \rightarrow V$, $x, y \in V$, we must have $d(k, x \vee_{k, g} y) \subseteq d(k, x) \cup d(k, y)$.
        \end{itemize}

@@

        We must now show that $\vee$ is commutative, associative, and idempotent, inductively. For any natural $n$ we say $\vee$ (applied to semilattice graphs) forms a semilattice up to $n$ if, for any $k : K$ whose longest dependency chain is no longer than $n$, we have:

        \begin{itemize}
          \item $(g_1 \vee g_2)(k) = (g_2 \vee g_1)(k)$
          \item $((g_1 \vee g_2) \vee g_3)(k) = (g_1 \vee (g_2 \vee g_3))(k)$
          \item $(g \vee g)(k) = g(k)$
        \end{itemize}

        We note that if $\vee$ forms a semilattice up to all $n$, then it forms a semilattice. We now note that if $\vee$ forms a semilattice up to $n$, then it forms a semilattice up to $n+1$, inductively:

        % WANT: (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k)
        %     = g_1(k) \vee_{k, g_1 \vee g_2 \vee g_3} (g_2(k) \vee_{k, g_2 \vee g_3} g_3(k))
        % HAVE: (g_1(k) \vee_{k, (g_1 \vee g_2 \vee g_3)}) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k)
        %     = g_1(k) \vee_{k, g_1 \vee g_2 \vee g_3} (g_2(k) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k))

        \begin{itemize}
          \item $(g_1 \vee g_2)(k) = g_2(k) \vee_{k, (g_1 \vee g_2)} g_1(k) = g_2(k) \vee_{k, (g_2 \vee g_1)} g_1(k) = (g_2 \vee g_1)(k)$
          \item
            \begin{align*}
              ((g_1 \vee g_2) \vee g_3))(k) &= (g_1 \vee g_2)(k) \vee_{k, ((g_1 \vee g_2) \vee g_3)} g_3(k) \\
              &= (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee (g_2 \vee g_3))} g_3(k) \\
              &= g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) \vee_{k, (g_1 \vee g_2)} g_3(k) \\
              &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2(k) \vee_{k, (g_1 \vee g_2)} g_3(k)) \\
              &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2 \vee g_3)(k) \\
              &= (g_1 \vee (g_2 \vee g_3))(k)
            \end{align*}
          \item $(g_1 \vee g_1)(k) = g_1(k) \vee_{k, (g_1 \vee g_1)} g_1(k) = g_1(k)$
        \end{itemize}
